Verifying Spatial Properties of Stencil Computations

### 2016-11-16 paper.tex
trimming in 5

Let $U$ range over the finite maps computed by \textit{modelise}.  For
each array variable $a \in \mathsf{dom}(U)$, the algorithm constructs next a set of
$n$-dimensional rectangles covering all contiguous groups of schemes
in $U(a)$.

@@

\begin{definition}[Spans]
  A \emph{span} represents an $n$-dimensional rectangle
  as a pair of $n$-dimensional vectors (or
  represented as a $2 \times n$ matrix) with values drawn from
  $\domainVal{}$. This gives the co-ordinates of a lower-bound vertex
  (first component) and the upper-bound (second component). We write
  spans as $\vect{x} = [\vect{x}^L \; \vect{x}^U]$, where $\vect{x}^L$
  and $\vect{x}^U$ give the lower and upper bounds.  In
  the following, $\vect{x}, \vect{y}$ range over spans.  For a span
  $\vect{x}$, we write $\vect{x}_{i:j}$ for the sub-span comprising the
  subvectors $[\vect{x}^L_{i:j} \; \vect{x}^U_{i:j}]$ from entry $i$
  to entry $j$ (inclusive).
\end{definition}

\noindent
Each indexing scheme $u \in U(a)$ is converted to a \emph{unit} span
$u \mapsto [u \; u]$, defining the lower and upper bound of an
$n$-dimensional cube of side length $1$.  For the five-point stencil
example there are five unit spans.
Figure~\ref{fig:inference-steps-informal}(a) illustrates these.

The next step repeatedly coalesces any contiguous spans until a fixed
point is reached, resulting in a set of (possibly overlapping) spans
covering the original space of schemes. Two $n$-dimensional spans
$\vect{x}$ and $\vect{y}$ are contiguous if for all but one
$j \in \{1, \ldots, n\}$ then $\vect{x}_j = \vect{y}_j$ and for one
$j$ then $\vect{x}^U_1 + 1 = \vect{y}^U$. For example, spans
$\left[\vtwoh{-1}{0}\; \vtwoh{-1}{0}\right]$ and
$[\vtwoh{0}{0} \; \vtwoh{1}{0}]$ are contiguous in the first dimension
and can be coalesced into the span $[\vtwoh{-1}{0} \; \vtwoh{1}{0}]$.

\begin{definition}\label{def:span-coalesc}
  Span coalescing $\bullet$ is defined for spans contiguous in the
  first dimension:
%%
\begin{align*}
\vect{x} \bullet \vect{y}
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n}) \\
\bot  & \textit{otherwise}
\end{cases}
% $m$-dimensions into coalesced spans which vary in
% $m+1$-dimensions.
\end{align*}
%%
\end{definition}
\noindent
To coalesce regions which are contiguous in other dimensions
other than the first, we generate all rotations of the spans in
$n$-dimensional space by permutation on spans, and fold together contiguous
regions using $\bullet$ (above). This
generates all possible ways of coalescing regions. For our example,
the unit spans in Figure~\ref{fig:inference-steps-informal}(a)
are coalesced into the contiguous spans in
Figure~\ref{fig:inference-steps-informal}(b).
This procedure is iterated until a fixed point is reached. In this
example this is reached in the first step.
 Then any spans that are completely contained by any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another), Figure~\ref{fig:inference-steps-informal}(c).
The final spans are used to synthesise concrete stencil specification
syntax (\S~\ref{sec:inf-step4}), where
each span becomes a product \texttt{*} of regions which are
combined by the sum \texttt{+}.

More formally the algorithm proceeds as follows.
For all $a \in \mathsf{dom}(U)$, let $M = U(a)$. Every scheme in $M$
is mapped to a unit span
by $\textsf{unitSpans}(M) = \{[u \; u] \mid u\!\in\!M\}$. For our example:
%%
\begin{align*}
  N(U(\texttt{a})) = \stwo{0}{0}{0}{0} \stwo{1}{0}{1}{0} \stwo{-1}{0}{-1}{0} \stwo{0}{1}{0}{1} \stwo{0}{-1}{0}{-1}
\end{align*}
%%
In the following we let $M_0 = N(U(\texttt{a}))$ for our example.

@@

\begin{figure}[t]
\vspace{-0.8em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
specification inference}
\label{fig:inference-steps-informal}
\vspace{-0.8em}
\end{figure}

\noindent
\textbf{1)\,} Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\begin{equation*}
P(M) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow M\}
\end{equation*}
%
This corresponds to all rotations of the space.  For our example:
%
\begin{align*}
P(M_0) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation flips the order of the two elements in the
vectors.

\noindent
\textbf{2)\,} Sort each permutation set into an ordered list, by the ordering:
\begin{equation*}
  \vect{x} \leq \vect{y} =
      \exists i \, . \; \vect{x}^L_{i} \leq \vect{y}^L_{i} \; \wedge \;
        (i = n \vee \vect{x}_{i+1:n} = \vect{y}_{i+1:n})
\end{equation*}
%
giving the sorting function $\textsf{sort}(M)$. For our example:
%%
\begin{align*}
\textsf{sort}(P(M_0)) =
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
\noindent
\textbf{3)\,} Reduce each list pairwise by $\bullet$ operation
(Def.~\ref{def:span-coalesc}) to coalesce contiguous regions,
given by $\textsf{fold}\bullet$. For our example, this yields:
%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}
]), (\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%
\textbf{4)\,} Un-permute and union together, \ie{},
%
\[
U(M) = \bigcup \{[\pi \vect{x}^L, \pi \vect{x}^U]
 \mid [\vect{x}^L, \vect{x}^U] \leftarrow S, (\pi, S) \leftarrow M\}
\]
For our example:
%
\begin{align*}
U(M) =
\{\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{-1}{0}{1},
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\textbf{5)\,} Filter by the region containment predicate $\containedin$, that
  is, if any region is contained within another then remove the
  smaller, defined:
%
\begin{align*}
& \vect{x} \containedin \vect{y} = \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{2:n}, \vect{x}^U_{2:n}) \containedin
  (\vect{y}^L_{2:n}, \vect{y}^U_{2:n})
\end{align*}
For our example this then yields the final result of:
\begin{align*}
\spanOp(M)
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$.
Applying \spanOp{} again yields the same
result for our example; the fixed point is reached.

@@

Abstract syntax trees of the specification language
are synthesised from the set $\spanOp(M)$ of coverings spans.
This conversion is defined in terms
of an algebra on \textit{spec} syntax in disjunctive-normal form
(which we denote \specDNF{}) and which is wrapped
in the $\textsf{Approx}$ type (see
Def.~\ref{def:mult-and-approx}).  The algebra mirrors the shape
of the region combinators for sum and product,
with operators $\oplus$ and $\odot$ and a constructor for \specDNF{}:
%%
\begin{align*}
\oplus & : \mathsf{Approx}(\specDNF) \rightarrow \mathsf{Approx}(\specDNF) \rightarrow
  \mathsf{Approx}(\specDNF) \\
\odot & : \mathsf{Approx}(\specDNF) \rightarrow \mathsf{Approx}(\specDNF) \rightarrow
        \mathsf{Approx}(\specDNF) \\
\mathsf{toSpec} & : \mathbb{N}_{>0} \rightarrow (\domainVal{} \times \domainVal{}) \rightarrow \mathsf{Approx}(\specDNF)
\end{align*}
%%
The $\mathsf{toSpec}$ function maps a dimension identifier
and a pair of lower and upper bound
values (one-dimensional) to a specification with a one-dimensional region.
An entire span is converted into an $n$-dimensional
specification by applying $\mathsf{toSpec}$ to each
pair of lower and upper bounds in a span, per dimension, and multiplying these
specifications by $\odot$. The resulting specifications for
each span are then summed by $\oplus$:
\begin{equation*}
\textsf{spec}(M) =
\oplusbig\limits_{\vect{x} \in \spanOp(M)} \,
\odobig\limits_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec} \; i \; (\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}
We unpack the definitions of this algebra.

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see Figure~\ref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec},
containing specifications in disjunctive normal form over regions
where the disjunct is \term{+} (since its model is in terms of set
union) and the conjunct is \term{*} (since its model is a form of tensor).
That is elements of \specDNF{} are of the form:
%%
$
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
$.

More formally, $\specDNF \subseteq (\textit{spec} \cup \{\epsilon\})$, defined:
%%
\begin{align*}
\specDNF{} & ::= \tySum{\specDNF} \mid \epsilon \\
\tySum{\specDNF} & ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
  \tyProd{\specDNF} \\
\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
   \textit{region}
\end{align*}
%%
We include the empty element $\epsilon$ which is not present in the
syntax.
In the following, $S, T$ range over elements of
\specDNF{} and $\tyProd{s}, \tyProd{t}$ range over elements
of $\tyProd{\specDNF}$ (for example, $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$).
\end{definition}

@@

\begin{figure}[t]
%\newcommand{\llower}{\vect{x}^L_i}
%\newcommand{\lupper}{\vect{x}^U_i}
\vspace{-0.7em}
\newcommand{\llower}{l}
\newcommand{\lupper}{u}
{\scalebox{0.97}{
\begin{minipage}{1\linewidth}
\begin{align*}
&\mathsf{toSpec} \; d \; (\llower, \lupper) = \\
&\hspace{-0.4em}\begin{cases}
%% POINT
\exact{\stenReflS{d}} & \llower = 0 \, \wedge \, \lupper = 0\\[0.2em]
%% FWD
\exact{\stenFwdS{\lupper}{d}}  & \llower = 0 \, \wedge \, \lupper > 0 \\[0.2em]
\exact{\stenFwdSR{\lupper}{d}{\irreflS}}\!\!\!\! & \llower = 1 \, \wedge \, \lupper > 0 \\[0.2em]
%% BWD
\exact{\stenBwdS{|\llower|}{d}}  & \llower < 0 \, \wedge \, \lupper = 0 \\[0.2em]
\exact{\stenBwdSR{|\llower|}{d}{\irreflS}}\!\!\!\! & \llower < 0 \, \wedge \, \lupper = -1 \\[0.2em]
%% CEN
\exact{\stenCenS{\lupper}{d}}& \llower < 0 \wedge \lupper > 0 \wedge |\llower| = \lupper \\[0.2em]
\exactR(\stenBwdS{|\llower|}{d} & \llower < 0 \wedge \lupper > 0  \wedge |\llower| \neq \lupper \\
\qquad\qquad \texttt{+} \, \stenFwdS{\lupper}{d}) & \\[0.2em]
% BOUND
\upperB{\stenFwdS{\lupper}{d}} & \llower > 1 \\[0.2em]
\upperB{\stenBwdS{|\llower|}{d}} \quad & \llower < (-1)
 \\[0.2em]
% ABSOLUTE REP
\exact{\epsilon} & \llower = \infty \, \vee \, \lupper = \infty
  \end{cases}
\end{align*}
\end{minipage}
}}
\caption{\textsf{toSpec} constructor of $\mathsf{Approx}(\specDNF)$ values.}
\label{fig:to-spec}
\vspace{-0.8em}
\end{figure}

\begin{definition} The $\mathsf{toSpec}$ constructor is defined
in Figure~\ref{fig:to-spec}.
The first two cases map to \texttt{forward} stencils since the lower
bound is $0$ or $1$, with the depth determined by the upper bound of
the span. If the lower bound is $1$, then the
$0$-offset point is missing so the region is given a
\texttt{nonpointed} attribute. The second two cases are similar for
\texttt{backward} but now the depth of the region is determined by the
absolute of the lower bound. In the fifth-case, where the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(sixth case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.

>>>>

Let $U$ range over the finite maps computed by \textit{modelise}.  For
each array variable $a \in \mathsf{dom}(U)$, the algorithm constructs a set of
$n$-dimensional rectangles covering all contiguous groups of schemes
in $U(a)$.

@@

\begin{definition}[Spans]
  A \emph{span} represents an $n$-dimensional rectangle
  as a pair of $n$-dimensional vectors (or
  represented as a $2 \times n$ matrix) with values drawn from
  $\domainVal{}$. This gives the co-ordinates of lower-bound 
  and upper-bound vertices. We write
  spans as $\vect{x} = [\vect{x}^L \; \vect{x}^U]$, where $\vect{x}^L$
  and $\vect{x}^U$ give the lower and upper bounds.  In
  the following, $\vect{x}, \vect{y}$ range over spans.  For a span
  $\vect{x}$, we write $\vect{x}_{i:j}$ for the sub-span comprising the
  subvectors $[\vect{x}^L_{i:j} \; \vect{x}^U_{i:j}]$ from entry $i$
  to entry $j$ (inclusive).
\end{definition}

\noindent
Each indexing scheme $u \in U(a)$ is converted to a \emph{unit} span
$u \mapsto [u \; u]$, defining the lower and upper bound of an
$n$-dimensional cube of side length $1$. For the five-point stencil
example there are five unit spans, illustrated in
Figure~\ref{fig:inference-steps-informal}(a).

The next step repeatedly coalesces any contiguous spans until a fixed
point is reached, resulting in a set of (possibly overlapping) spans
covering the original space of schemes. Two $n$-dimensional spans
$\vect{x}$ and $\vect{y}$ are contiguous if for all but one
$j \in \{1, \ldots, n\}$ then $\vect{x}_j = \vect{y}_j$ and for one
$j$ then $\vect{x}^U_1 + 1 = \vect{y}^L$. For example, spans
$\left[\vtwoh{-1}{0}\; \vtwoh{-1}{0}\right]$ and
$[\vtwoh{0}{0} \; \vtwoh{1}{0}]$ are contiguous in the first dimension
and can be coalesced into the span $[\vtwoh{-1}{0} \; \vtwoh{1}{0}]$.

\begin{definition}\label{def:span-coalesc}
  Span coalescing $\bullet$ is partial, defined for spans contiguous in the
  first dimension:
%%
\begin{align*}
\vect{x} \bullet \vect{y}
= (\vect{x}^L, \vect{y}^U) \;\;\;\, \textit{if} \;\;\; \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n} %\\
%\bot  & \textit{otherwise}
%\end{cases}
% $m$-dimensions into coalesced spans which vary in
% $m+1$-dimensions.
\end{align*}
%%
\end{definition}
\noindent
To coalesce regions which are contiguous in other dimensions
other than the first, we generate all rotations of the spans in
$n$-dimensional space by permutation on spans, and fold together contiguous
regions using $\bullet$ (above). This
generates all possible ways of coalescing regions. For our example,
the unit spans in Figure~\ref{fig:inference-steps-informal}(a)
are coalesced into the contiguous spans in
Figure~\ref{fig:inference-steps-informal}(b).
This procedure is iterated until a fixed point is reached (in one step
in this example).
 Then any spans that are completely contained by any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another), Figure~\ref{fig:inference-steps-informal}(c).
The final spans are used to synthesise concrete stencil specification
syntax (\S~\ref{sec:inf-step4}), where
each span becomes a product \texttt{*} of regions which are
combined by the sum \texttt{+}.

More formally the algorithm proceeds as follows.
For all $a \in \mathsf{dom}(U)$, let $M = U(a)$. Every scheme in $M$
is mapped to a unit span
by $\textsf{unitSpans}(M) = \{[u \; u] \mid u\!\in\!M\}$, \eg{}:
%%
\begin{align*}
  N(U(\texttt{a})) = \stwo{0}{0}{0}{0} \stwo{1}{0}{1}{0} \stwo{-1}{0}{-1}{0} \stwo{0}{1}{0}{1} \stwo{0}{-1}{0}{-1}
\end{align*}
%%
In the following we let $M_0 = N(U(\texttt{a}))$ for our example.

@@

\begin{figure}[t]
\vspace{-0.8em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
inference}
\label{fig:inference-steps-informal}
\vspace{-0.8em}
\end{figure}

\noindent
\textbf{1)\,} Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\begin{equation*}
P(M) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow M\}
\end{equation*}
%
This corresponds to all rotations of the space, \eg{}:
%
\begin{align*}
P(M_0) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation that flips the order of the two elements in the
vectors.

\noindent
\textbf{2)\,} Sort each permutation set into a list, by the ordering:
\begin{equation*}
  \vect{x} \leq \vect{y} =
      \exists i \, . \; \vect{x}^L_{i} \leq \vect{y}^L_{i} \; \wedge \;
        (i = n \vee \vect{x}_{i+1:n} = \vect{y}_{i+1:n})
\end{equation*}
%
giving the sorting function $\textsf{sort}(M)$, \eg{}, $\textsf{sort}(P(M_0))$:
%%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
\noindent
\textbf{3)\,} Reduce each list pairwise by $\bullet$
(Def.~\ref{def:span-coalesc}) to coalesce contiguous regions,
given by $\textsf{fold}\bullet$. For our example:
%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}
]), (\pi^2_2, \textit{same as for $\pi_1$})\}
\end{align*}
%
\textbf{4)\,} Un-permute and union together, \ie{},
%
\[
U(M) = \bigcup \{[\pi \vect{x}^L, \pi \vect{x}^U]
 \mid [\vect{x}^L, \vect{x}^U] \leftarrow S, (\pi, S) \leftarrow M\}
\]
For our example:
%
\begin{align*}
U(M) =
\{\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{-1}{0}{1},
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\textbf{5)\,} Filter by the region containment predicate
$\containedin$; if any region is contained within another then remove the
  smaller, defined:
%
\begin{align*}
& \vect{x} \containedin \vect{y} = \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{2:n}, \vect{x}^U_{2:n}) \containedin
  (\vect{y}^L_{2:n}, \vect{y}^U_{2:n})
\end{align*}
For our example this then yields the final result of:
\begin{align*}
\spanOp(M)
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$.
Applying \spanOp{} again yields the same
result for our example.%; the fixed point is reached.

@@

Abstract syntax trees of the specification language
are synthesised from the set $\spanOp(M)$ of coverings spans.
This conversion is defined in terms
of an algebra on \textit{spec} syntax in disjunctive-normal form
(which we denote \specDNF{}) and which is wrapped
in the $\textsf{Approx}$ type (see
Def.~\ref{def:mult-and-approx}).  The algebra mirrors the shape
of the region operators $\term{+}$ and $\term{*}$ 
with binary operators $\oplus$ and $\odot$ on
$\mathsf{Approx}(\specDNF)$.
A function $\mathsf{toSpec}$ maps a dimension identifier
and a pair of lower and upper bound
values (one-dimensional) to a specification with a one-dimensional region.
%%
\begin{align*}
\mathsf{toSpec} & : \mathbb{N}_{>0} \rightarrow (\domainVal{} \times \domainVal{}) \rightarrow \mathsf{Approx}(\specDNF)
\end{align*}
%%
An entire span is converted into an $n$-dimensional
specification by applying $\mathsf{toSpec}$ to each
pair of lower and upper bounds in a span, per dimension, and multiplying these
specifications by $\odot$. The resulting specifications for
each span are then summed by $\oplus$:
\begin{equation*}
\textsf{spec}(M) =
\oplusbig\limits_{\vect{x} \in \spanOp(M)} \,
\odobig\limits_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec} \; i \; (\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}
We unpack the definitions of this algebra.

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see Fig.~\ref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec},
containing specifications in disjunctive normal form over regions
where the disjunct is \term{+} and the conjunct is \term{*}; 
 elements of \specDNF{} are of the form:
%%
$
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
$.

More formally, $\specDNF \subseteq (\textit{spec} \cup \{\epsilon\})$, defined:
%%
\begin{align*}
\specDNF{} & ::= \tySum{\specDNF} \mid \epsilon \quad\;\;
\tySum{\specDNF} ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
  \tyProd{\specDNF} \\
\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
   \textit{region}
\end{align*}
%%
In the following, $S, T$ range over elements of
\specDNF{} and $\tyProd{s}, \tyProd{t}$ range over elements
of $\tyProd{\specDNF}$, \ie{} $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$.
\end{definition}

@@

\begin{figure}[t]
%\newcommand{\llower}{\vect{x}^L_i}
%\newcommand{\lupper}{\vect{x}^U_i}
\vspace{-0.7em}
\newcommand{\llower}{l}
\newcommand{\lupper}{u}
{\scalebox{0.97}{
\begin{minipage}{1\linewidth}
\begin{align*}
&\mathsf{toSpec} \; d \; (\llower, \lupper) = \\
&\hspace{-0.4em}\begin{cases}
%% POINT
\exact{\stenReflS{d}} & \llower = 0 \, \wedge \, \lupper = 0\\[0.02em]
%% FWD
\exact{\stenFwdS{\lupper}{d}}  & \llower = 0 \, \wedge \, \lupper > 0 \\[0.02em]
\exact{\stenFwdSR{\lupper}{d}{\irreflS}}\!\!\!\! & \llower = 1 \, \wedge \, \lupper > 0 \\[0.02em]
%% BWD
\exact{\stenBwdS{|\llower|}{d}}  & \llower < 0 \, \wedge \, \lupper = 0 \\[0.02em]
\exact{\stenBwdSR{|\llower|}{d}{\irreflS}}\!\!\!\! & \llower < 0 \, \wedge \, \lupper = -1 \\[0.02em]
%% CEN
\exact{\stenCenS{\lupper}{d}}& \llower < 0 \wedge \lupper > 0 \wedge |\llower| = \lupper \\[0.02em]
\exactR(\stenBwdS{|\llower|}{d} & \llower < 0 \wedge \lupper > 0  \wedge |\llower| \neq \lupper \\
\qquad\qquad \texttt{+} \, \stenFwdS{\lupper}{d}) & \\[0.02em]
% BOUND
\upperB{\stenFwdS{\lupper}{d}} & \llower > 1 \\[0.02em]
\upperB{\stenBwdS{|\llower|}{d}} \quad & \llower < (-1)
 \\[0.02em]
% ABSOLUTE REP
\exact{\epsilon} & \llower = \infty \, \vee \, \lupper = \infty
  \end{cases}
\end{align*}
\end{minipage}
}}
\caption{\textsf{toSpec} constructor of $\mathsf{Approx}(\specDNF)$ values.}
\label{fig:to-spec}
\vspace{-0.8em}
\end{figure}

\begin{definition} The $\mathsf{toSpec}$ constructor is defined
in Figure~\ref{fig:to-spec}, where the lower and upper bounds
determine region constants; $\mathsf{toSpec} \, d \, (l, u)$ is always defined
 if $l$ less than or equal to $u$.

Positive lower and upper bound yield \term{forward} regions, and
negative lower and upper bounds yield \term{backward} regions.
The \texttt{nonpointed} attribute (abbreviated \term{np})
is present when the lower bound is at $1$ or $-1$. 
In the sixth-case, where the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(seventh case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.
If the lower bound is greater than 1
(which implies the upper bound is greater than 1 too) then
this represents a region that is not at the origin (or next to) the
origin and thus cannot be represented in the specification language.
Thus, an upper bound is produced showing the maximum extent
of the region. The penultimate case is the dual, when the lower bound is less
than $-1$, giving an upper bound on a \texttt{backward} specification.
\end{definition}
